Borel regularity for ODEs and Thimble Integrals

### 2023-12-21 exp-resurgence.tex
Updates from Overleaf

\item\label{part-1} As $z \to \infty$ along the ray $z \in e^{-i\theta}\,\R_{>0}$, the function $I_j$ is asymptotic to a trans-monomial $\tilde{I}_{j}\in e^{-z \alpha_j} z^{-1/2} \C\llbracket z^{-1}\rrbracket$.

@@

Let's write $\approx$ when two functions are asymptotic at all orders around the base poin; by the steepest descent method,\footnote{The details can be found in \cite{miller2006applied}: follow the proof of Proposition~2.1 in through equation~(2.9).} t  
%%\[ e^{-z\zeta_\alpha} I_\alpha(z) \approx \int_{\tau \in [-\varepsilon, \varepsilon]} e^{-z\tau^2/2} \nu \]
\[ e^{z \alpha_j} I_j \approx \int_{\tau \in [-\varepsilon, \varepsilon]} e^{-z\tau^2/2} \nu \]
as $z \to \infty$ along the ray $e^{-i\theta}\, \R_{> 0}$. Plugging in the Taylor series for $\nu$, we get
\begin{align*}
e^{z \alpha_j} I_j & \approx \int_{-\varepsilon}^\varepsilon e^{-z\tau^2/2} \sum_{m \ge 0} b_m^j \tau^m\,d\tau \\
& = \int_{-\varepsilon}^\varepsilon e^{-z\tau^2/2} \sum_{n \ge 0} b_{2n}^j \tau^{2n}\,d\tau.
\end{align*}
By the dominated convergence theorem,\footnote{Notice that the sum over $k$ is empty when $n = 0$. Following convention, we extend the double factorial to all odd integers by its recurrence relation, giving $(-1)!! = 1$.}
\begin{align*}
e^{z \alpha_j} I_j & \approx \sum_{n \ge 0} b_{2n}^j \int_{-\varepsilon}^\varepsilon e^{-z\tau^2/2} \tau^{2n}\,d\tau \\
& = \sum_{n \ge 0} (2n-1)!!\,b_{2n}^j \left[ \sqrt{2\pi}\,z^{-(n+1/2)} \operatorname{erf}\big(\varepsilon \sqrt{z/2}\big) - 2e^{-z\varepsilon^2/2} \sum_{\substack{k \in \mathbb{N}_+ \\ k \le n}} \frac{\varepsilon^{2k-1}}{(2k-1)!!} z^{-n+k-1} \right].
\end{align*}

@@

Rewriting equation~\eqref{thimble-difference} as
\[ \hat{\iota}_j(\zeta) = \partial_\zeta \left( \int_{\Lambda_j(\zeta)} \nu \right), \]
we can see that
\[ \fracderiv{-1}{\zeta}{\alpha_j} \hat{\iota}_j(\zeta) = \int_{\Lambda_j(\zeta)} \nu - \int_{\Lambda_j(\alpha_j)} \nu. \]
The initial value term vanishes because the path $\alpha_j$ is a point. Hence,
\[ \int_{\Lambda_j(\zeta)} \nu = \fracderiv{-3/2}{\zeta}{\alpha_j} \hat{\phi}_j(\zeta). \]
Recalling that the Riemann-Liouville fractional derivative is a left inverse of the fractional integral, we conclude that
\[ \fracderiv{3/2}{\zeta}{\alpha_j} \left( \int_{\Lambda_j(\zeta)} \nu \right) = \hat{\phi}_j(\zeta). \]
\end{proof}
\color{black}
\begin{remark}
    The $3/2$ derivative formula~\eqref{formula1} allows us to compute the sum of the germ $\series{\phi}_j$, hence it is enough to use it to study the location of singularities in the position domain and its growth behavior at infinity\footnote{For instance in \cite{sauzin2021variations}, the analogous of formula~\eqref{formula1} is used to compute a close formula for the Borel sum of the Stirling approximation of the Gamma function.}. Furthermore, expanding $\hat{\phi}_j$ at its critical points, we recover the germs $\series{\phi}_k$ for $k\neq j$. This is an instance of resurgence, indeed according to the theory of \'Ecalle, $\hat{\phi}_j$ are all resurgent functions. We'll see an example in Appendix~\ref{apx:eye-res-airy}. 
\end{remark}
\textcolor{orange}{
\begin{corollary}\label{simple-res-thimble}
    If $f$ and $\nu$ are polynomials, the function $\hat{\phi}_j$ defined in \eqref{formula1} has simple resurgent singularities at $\zeta=\alpha_k$.  
\end{corollary}
\begin{proof}
    Recall $\hat{\iota}_j$ is the sum of the Borel transform of $z^{1/2}\tilde{I}_j$, namely 
    \[\hat{\iota}_j(\zeta)=\fracderiv{1/2}{\zeta}{\alpha_j}\hat{\iota}_j=\int_{\alpha_j}^\zeta(\zeta-s)^{-3/2}\hat{\iota}_j(s)\, ds\]
    Then by Pham's result (see \eqref{rmk:Pham formula}) we know 
    \[\hat{\iota}_j(\zeta_j)=\frac{(-\zeta_j)^{-1/2}}{2\pi i\Gamma(-\tfrac{1}{2})}\sum_{k=0}^\infty c_k (-\zeta_j)^{-k} + \text{ hol. funct. }\]
hence 
\[\hat{\iota}_j(\zeta)=\int_{\alpha_j}^\zeta(\zeta-s)^{-3/2}\frac{(-s+\alpha_j)^{-1/2}}{2\pi i\Gamma(-\tfrac{1}{2})}\sum_{k=0}^\infty c_k (-s+\alpha_j)^{-k} \, ds + \text{ hol. funct. }\]
\end{proof}
}
\begin{remark}\label{rmk:Pham formula}
    When $f$ and $\nu$ are polynomials and $X$ is $N$-dimensional, a result of Pham \cite[Equation 2.4, II partie]{pham} allows to write ${\iota}_j$ explicitly as 
    \begin{equation}\label{eqn:Pham}
        (-1)^{\frac{N(N-1)}{2}}  \sum_{k=0}^\infty c_k\, \delta^{-\frac{N}{2}-k}+ \text{ hol. funct.}
    \end{equation}
    for some coefficients $c_k$ which depends on $f$ and $\nu$, and with $\delta^{-\alpha}$ be defined as
   \begin{equation*}
       \delta^{-\alpha}:=\begin{cases}
           \frac{\zeta^{\alpha-1}}{-2\pi i(\alpha-1)!}\log(\zeta) + \text{ hol. funct.} & \text{ if } \alpha\in\mathbb{N}^*\\
           & \\
           \frac{(-\zeta)^{\alpha-1}}{2\pi i \Gamma(-\alpha)}+ \text{ hol. funct.} & \text{ if } \alpha\notin \mathbb{N}^* 
       \end{cases}
   \end{equation*}
In particular, if $N=1$, from equation \eqref{eqn:Pham} we deduce that $\hat{\phi}_j$ is the $1/2-$ derivative of ${\iota}_j$, \textcolor{orange}{hence it has simple singularities \ref{simple-res-thimble}.} We expect this result to hold beyond the polynomial case. We also expect that equation \eqref{eqn:formula} can be generalized to the $N$-dimensional case replacing the $3/2$-derivative with the $(N+1)/2$ one. 
\end{remark}
\section{Examples}\label{sec:examples}

>>>>

\item\label{part-1} As $z \to \infty$ along the ray $z \in e^{-i\theta}\,\R_{>0}$, the function $I_j$ is asymptotic to a trans-monomial $\tilde{I}_{j}\in e^{-z \alpha_j} z^{-1/2} \C\llbracket z^{-1}\rrbracket$. Recall that $\theta$ is the direction of the ray $\Gamma^\theta_j$.

@@

Let's write $\approx$ when two functions are asymptotic at all orders around the base poin; by the steepest descent method,\footnote{The details can be found in \cite{miller2006applied}: follow the proof of Proposition~2.1 in through equation~(2.9).}
%%\[ e^{-z\zeta_\alpha} I_\alpha(z) \approx \int_{\tau \in [-\varepsilon, \varepsilon]} e^{-z\tau^2/2} \nu \]
\[ e^{z \alpha_j} I_j \approx \int_{\tau \in [-\varepsilon, \varepsilon]} e^{-z\tau^2/2} \nu \]
as $z \to \infty$ along the ray $e^{-i\theta}\, \R_{> 0}$. Plugging in the Taylor series for $\nu$, we get
\begin{align*}
e^{z \alpha_j} I_j & \approx \int_{-\varepsilon}^\varepsilon e^{-z\tau^2/2} \sum_{m \ge 0} b_m^j \tau^m\,d\tau \\
& = \int_{-\varepsilon}^\varepsilon e^{-z\tau^2/2} \sum_{n \ge 0} b_{2n}^j \tau^{2n}\,d\tau.
\end{align*}
By the dominated convergence theorem,\footnote{Notice that the sum over $k$ is empty when $n = 0$. Following convention, we extend the double factorial to all odd integers by its recurrence relation, giving $(-1)!! = 1$.}
\begin{align*}
e^{z \alpha_j} I_j & \approx \sum_{n \ge 0} b_{2n}^j \int_{-\varepsilon}^\varepsilon e^{-z\tau^2/2} \tau^{2n}\,d\tau \\
& = \sum_{n \ge 0} (2n-1)!!\,b_{2n}^j \left[ \sqrt{2\pi}\,z^{-(n+1/2)} \operatorname{erf}\big(\varepsilon \sqrt{z/2}\big) - 2e^{-z\varepsilon^2/2} \sum_{\substack{k \in \mathbb{N}_+ \\ k \le n}} \frac{\varepsilon^{2k-1}}{(2k-1)!!} z^{-n+k-1} \right].
\end{align*}

@@

Rewriting equation~\eqref{thimble-difference} as
\[ \hat{\iota}_j(\zeta) = \partial_\zeta \left( \int_{\Lambda_j(\zeta)} \nu \right), \]
we can see that
\[ \fracderiv{-1}{\zeta}{\alpha_j} \hat{\iota}_j(\zeta) = \int_{\Lambda_j(\zeta)} \nu - \int_{\Lambda_j(\alpha_j)} \nu. \]
The initial value term vanishes because the path $\alpha_j$ is a point. Hence,
\[ \int_{\Lambda_j(\zeta)} \nu = \fracderiv{-3/2}{\zeta}{\alpha_j} \hat{\phi}_j(\zeta). \]
Recalling that the Riemann-Liouville fractional derivative is a left inverse of the fractional integral, we conclude that
\[ \fracderiv{3/2}{\zeta}{\alpha_j} \left( \int_{\Lambda_j(\zeta)} \nu \right) = \hat{\phi}_j(\zeta). \]
\end{proof}
\color{black}
\begin{remark}
We can use the $3/2$ derivative formula~\eqref{formula1} to see that the functions $\phi_j$ are all resurgent, in the sense of \'{E}calle~\cite[\textcolor{orange}{section?}]{EcalleI}. Let's say $\Gamma_j(\beta)$ is the straight path from $\zeta = \alpha_j$ to $\zeta = \beta$ in $\C$. As long as this path avoids the critical values of $f$, it lifts uniquely along $f$ to a path $\Lambda_j(\beta)$ in $X$. This lets us analytically continue $\int_{\Lambda_j(\zeta)} \nu$ to a star-shaped domain $\Omega \subset \C$. Intuitively, $\Omega$ is constructed by drawing rays from $\zeta = \alpha_j$ through all of the other critical values, and then cutting $\C$ along the parts of those rays that go from the other critical values to $\infty$.
\begin{center}
\textcolor{orange}{FIGURE OF SLIT DOMAIN}
\end{center}
Let's calculate the jump of $\int_{\Lambda_j(\zeta)}$ across the cut starting at $\zeta = \alpha_k$. Write $\alpha_k$ as $\alpha_j + re^{i\theta}$, and let $\alpha_k^+$ and $\alpha_k^-$ be the points we get by infinitesimally increasing or decreasing $\theta$. The jump is then given by $\int_{\Lambda_j(\alpha_k^+)} \nu - \int_{\Lambda_j(\alpha_k^-)} \nu$.

\textcolor{Navy}{Initially, the function $\int_{\Lambda_j(\zeta)} \nu$ is only defined for $\zeta$ on the ray $[\alpha_j, \infty)$, but it can be analytically continued to points off the ray by homotopy of the path $\Lambda_j(\zeta)$. Since $\nu$ is holomorphic, the only way for $\int_{\Lambda_j(\zeta)} \nu$ to be singular is for something to go wrong with the hotmotopy of $\Lambda_j(\zeta)$. At a critical point of $f$, the ends of the homotoped path $\Lambda_j(\zeta)$ come together, telling us that.}
\end{remark}
\color{DarkSeaGreen}
\begin{remark}
    The $3/2$ derivative formula~\eqref{formula1} allows us to compute the sum of the germ $\series{\phi}_j$, hence it is enough to use it to study the location of singularities in the position domain and its growth behavior at infinity\footnote{For instance in \cite{sauzin2021variations}, the analogous of formula~\eqref{formula1} is used to compute a close formula for the Borel sum of the Stirling approximation of the Gamma function.}. Furthermore, expanding $\hat{\phi}_j$ at its critical points, we recover the germs $\series{\phi}_k$ for $k\neq j$. This is an instance of resurgence, indeed according to the theory of \'{E}calle, $\hat{\phi}_j$ are all resurgent functions. We'll see an example in Appendix~\ref{apx:eye-res-airy}. 
\end{remark}
\color{black}
\textcolor{orange}{
\begin{corollary}\label{simple-res-thimble}
    If $f$ and $\nu$ are polynomials, the function $\hat{\phi}_j$ defined in \eqref{formula1} has simple resurgent singularities at $\zeta=\alpha_k$.  
\end{corollary}
\begin{proof}
    Recall $\hat{\iota}_j$ is the sum of the Borel transform of $z^{1/2}\tilde{I}_j$, namely 
    \[\hat{\iota}_j(\zeta)=\fracderiv{1/2}{\zeta}{\alpha_j}\hat{\iota}_j=\int_{\alpha_j}^\zeta(\zeta-s)^{-3/2}\hat{\iota}_j(s)\, ds\]
    Then by Pham's result (see \eqref{rmk:Pham formula}) we know 
    \[\hat{\iota}_j(\zeta_j)=\frac{(-\zeta_j)^{-1/2}}{2\pi i\Gamma(-\tfrac{1}{2})}\sum_{k=0}^\infty c_k (-\zeta_j)^{-k} + \text{ hol. funct. }\]
hence 
\[\hat{\iota}_j(\zeta)=\int_{\alpha_j}^\zeta(\zeta-s)^{-3/2}\frac{(-s+\alpha_j)^{-1/2}}{2\pi i\Gamma(-\tfrac{1}{2})}\sum_{k=0}^\infty c_k (-s+\alpha_j)^{-k} \, ds + \text{ hol. funct. }\]
\end{proof}
}
\begin{remark}\label{rmk:Pham formula}
    When $f$ and $\nu$ are polynomials and $X$ is $N$-dimensional, a result of Pham \cite[Equation 2.4, II partie]{pham} allows to write ${\iota}_j$ explicitly as 
    \begin{equation}\label{eqn:Pham}
        (-1)^{\frac{N(N-1)}{2}}  \sum_{k=0}^\infty c_k\, \delta^{-\frac{N}{2}-k}+ \text{ hol. funct.}
    \end{equation}
    for some coefficients $c_k$ which depends on $f$ and $\nu$, and with $\delta^{-\alpha}$ be defined as
   \begin{equation*}
       \delta^{-\alpha}:=\begin{cases}
           \frac{\zeta^{\alpha-1}}{-2\pi i(\alpha-1)!}\log(\zeta) + \text{ hol. funct.} & \text{ if } \alpha\in\mathbb{N}^*\\
           & \\
           \frac{(-\zeta)^{\alpha-1}}{2\pi i \Gamma(-\alpha)}+ \text{ hol. funct.} & \text{ if } \alpha\notin \mathbb{N}^* 
       \end{cases}
   \end{equation*}
In particular, if $N=1$, from equation \eqref{eqn:Pham} we deduce that $\hat{\phi}_j$ is the $1/2-$ derivative of ${\iota}_j$, \textcolor{orange}{hence it has simple singularities \ref{simple-res-thimble}.} We expect this result to hold beyond the polynomial case. We also expect that equation \eqref{eqn:formula} can be generalized to the $N$-dimensional case replacing the $3/2$-derivative with the $(N+1)/2$ one. 
\end{remark}
\section{Examples}\label{sec:examples}
